---
title: "Synthetic data"
author: "Stijn van den Broek"
date: "02/05/2021"
output:
  html_document: default
  pdf_document: default
---

```{r, results='hide', message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(magrittr)
library(corrplot)
library(mice)
library(furrr)
library(purrr)
library(broom)
library(caret)
library(mboost)
library(e1071)
library(xtable)
library(DT)
library(psych)
```

## Setting the seed

```{r}
set.seed(123)
```

## Reading the data

```{r}
data = read.csv('diabetes.csv')
data$Outcome = factor(data$Outcome, levels = c(0, 1))
```

## Creating a version of the data where the missing data is imputed with a low amount of iterations and a high amount of iterations

```{r}
imputed_data = data

imputed_data[,2:8][imputed_data[,2:8] == 0] = NA #setting all zeroes from all columns apart from pregnancies and outcome to NA

print(md.pattern(imputed_data))

imputed_data = mice(imputed_data, m=5, maxit=5, meth='pmm', seed=123)
imputed_data = complete(imputed_data, 1)
```

# Functions

The following functions are used throughout the document.

## Function for obtaining statistical properties

```{r}
statistical_properties = function(dataframe){
  stats = data.frame()
  for (i in seq(1, length(colnames(dataframe)), 1)) {
    column_data = as.numeric(dataframe[,i])
    term = colnames(dataframe)[i]
    mean = mean(column_data)
    standard_deviation = sd(column_data)
    skewness = skewness(column_data)
    kurtosis = kurtosis(column_data)
    standard_error = sd(column_data)/sqrt(length(column_data))
    row = data.frame(term = term, mean = mean, standard_deviation = standard_deviation, skewness = skewness, kurtosis = kurtosis, standard_error = standard_error)
    stats = rbind(stats, row)
  }
  return(stats)
}
```

## Function for plotting distributions

```{r}
dataset_distribution = function(dataframe, bins = 20) {
  ggplot(gather(dataframe), aes(as.numeric(value))) + 
    geom_histogram(bins = bins) + 
    facet_wrap(~key, scales = 'free_x')
}
```

## Synthesizing function and imputation methods

```{r}
synthesize = function(dataframe, method = "cart", bootstrap = FALSE, m = 5, nsim = 100, maxit = 1, cp = 0.0001) {
  set.seed(123)
  if (bootstrap == TRUE){
    syn <- future_map(1:nsim, function(x){
                  mice(dataframe[sample(1:768, 768, replace = TRUE), ], 
                          m=m, 
                          meth = method, 
                          where = matrix(TRUE, nrow(dataframe), ncol(dataframe)), 
                          cp = cp,
                          minbucket = 3,
                          print = FALSE)}, 
                  .options = furrr_options(seed = as.integer(123)), .progress = T)
  }
  
  else {
    syn <- future_map(1:nsim, function(x){
                  mice(dataframe, 
                          m=m, 
                          meth = method, 
                          where = matrix(TRUE, nrow(dataframe), ncol(dataframe)), 
                          cp = cp,
                          minbucket = 3,
                          print = FALSE)}, 
                  .options = furrr_options(seed = as.integer(123)), .progress = T)
  }
  
  return(syn)
}
```

## Pooling function

```{r}
pool3.syn <- function(mira) {
  set.seed(123)
  if(class(mira)[1] == "mira") { # if the input object is of class mira
    fitlist <- mira %$% analyses # extract the analyses from the mira object
  }
  else {
    fitlist <- mira              # and otherwise, just take the input list
  }
  
  vars <- fitlist[[1]] %>% coef() %>% names()
  
  m <- length(fitlist)           # number of imputations
  
  pooled <- fitlist %>% 
    map_dfr(broom::tidy) %>%
    group_by(term) %>%
    summarise(est     = mean(estimate),
              bm      = sum((estimate - est)^2) / (m - 1),
              ubar    = mean(std.error^2),
              var     = ubar + bm/m, # new variance estimate
              df      = (m - 1) * (1 + (ubar * m)/bm), # and new df estimate
              lower   = est - qt(.975, df) * sqrt(var),
              upper   = est + qt(.975, df) * sqrt(var), .groups = 'drop') %>%
    arrange(factor(term, levels = vars))
  pooled
}
```

## Test pooling function for bootstrapped data

```{r}
pool.normal <- function(mira) {
  
  if(class(mira)[1] == "mira") { # if the input object is of class mira
    fitlist <- mira %$% analyses # extract the analyses from the mira object
  }
  else {                         # and otherwise, just take the input list
    fitlist <- mira
  }
  
  vars <- fitlist[[1]] %>% coef() %>% names()
  
  m <- length(fitlist)           # number of imputations
  
  pooled <- fitlist %>% 
    map_dfr(broom::tidy) %>%     # tidy estimates
    group_by(term) %>%           # group per variable
    summarise(est     = mean(estimate),
              bm      = sum((estimate - est)^2) / (m - 1),
              ubar    = mean(std.error^2),
              var_u   = (1 + 1/m) * bm + ubar,
              var     = if_else(var_u > 0, var_u, ubar), # restrict variance to be positive
              df      = max(1, (m - 1) * (1 - ubar / (bm + bm/m))^2), # restrict df > 1
              lower   = est - qt(.975, df) * sqrt(var),
              upper   = est + qt(.975, df) * sqrt(var), .groups = 'drop') %>%
    arrange(factor(term, levels = vars))
  pooled
}
```

## Confidence interval function

```{r}
ci_cov <- function(pooled, true_fit = NULL, coefs = NULL, vars = NULL) {
  set.seed(123)
  
  if (!is.null(true_fit)) {
    coefs <- coef(true_fit)
    vars   <- diag(vcov(true_fit))
  }
  
  nsim <- nrow(pooled) / length(unique(pooled$term))
  
  pooled %>% mutate(true_coef = rep(coefs, times = nsim),
                    true_var  = rep(vars, times = nsim),
                    cover     = lower < true_coef & true_coef < upper) %>%
    group_by(term) %>%
    summarise("True Est" = unique(true_coef),
              "Syn Est"  = mean(est),
              "Bias"     = mean(est - true_coef),
              "True SE"  = unique(sqrt(true_var)),
              "Syn SE"   = mean(sqrt(var)),
              "df"       = mean(df),
              "Lower"    = mean(lower),
              "Upper"    = mean(upper),
              "CIW"      = mean(upper - lower),
              "Coverage" = mean(cover), .groups = "drop")
}
```

## Prediction function

```{r}
predict_synth_or_true = function(sim, origin) {
  truemodel <- origin %$%
  glm(Outcome ~ BMI + Glucose + Pregnancies, family = "binomial") #model
  
  sim %>%
    map(~.x %$% 
          glm(Outcome ~ BMI + Glucose + Pregnancies, family = "binomial")) %>%
    map_dfr(pool3.syn) %>%
    ci_cov(., truemodel) %>% 
    datatable()
  
  out <- sim %>%
    map(~.x %>% 
          complete("all") %>%  
          map(~.x %>% 
                rbind(origin) %>% 
                mutate(indic = as.factor(c(rep(0, 768), rep(1, 768)))) %$% 
                glm(Outcome ~ BMI + Glucose + Pregnancies, family = "binomial")) %>% pool() %>%  summary() %>% as.data.frame() %>% column_to_rownames(var = "term")) %>% 
    Reduce("+", .) %>% round(digits = 4)
  datatable(out / 100)
  
  do_fun <- function(x){
    dat <- x %>% 
      rbind(origin) %>% 
      mutate(indic = as.factor(c(rep(0, 768), rep(1, 768)))) 
    fit <- dat %$% 
      glm(Outcome ~ BMI + Glucose + Pregnancies, family = "binomial") %>% 
      predict(type = "response")
    CM <- confusionMatrix(factor(as.numeric(fit > .5)), factor(dat$indic))
    c(CM$overall, CM$byClass)
  }
  sim %>%
    map(~.x %>% 
          complete("all") %>%  
          map(~.x %>% do_fun()) %>% 
          do.call("rbind", .) %>% colMeans()) %>% 
    do.call("rbind", .) %>% colMeans() %>% .[c(1, 2, 8, 9, 15, 18)] %>% round(digits = 4)
}
```

# The process of synthesizing

These are the steps for synthesizing and analyzing the data.

## 1. Formulating an analysis model

```{r}
analysis_model = glm(Outcome~BMI+Glucose+Pregnancies, data = data, family = "binomial")
```

## 2. Running the analysis model on the data set to obtain the true data inference

```{r}
summary(analysis_model)
```

## 3. Obtaining the statistical properties of the true data set

```{r}
true_statistics = statistical_properties(data)
```

## 4. Plotting the distributions of the true data set

```{r, warning=FALSE}
dataset_distribution(data)
dataset_distribution(imputed_data)
```

## 5a. Synthesizing the data

```{r, warning=FALSE}
plan(multisession)

synthesized = synthesize(data, nsim = 1000) #the standard synthesis with original data, no bootstrap and a high cp
ori_nb_lcp = synthesize(data, nsim = 1000, cp = 1e-32) #synthesis with original data, no bootstrap and a low cp
orl_b_hcp = synthesize(data, nsim = 1000, bootstrap = TRUE) #synthesis with original data, bootstrap, low iterations and a high cp
orl_b_lcp = synthesize(data, nsim = 1000, bootstrap = TRUE, cp = 1e-32) #synthesis with original data, bootstrap, low iterations and a low cp
orh_b_hcp = synthesize(data, nsim = 1000, bootstrap = TRUE, maxit = 5) #synthesis with original data, bootstrap, high iterations and a high cp
orh_b_lcp = synthesize(data, nsim = 1000, bootstrap = TRUE, cp = 1e-32, maxit = 5) #synthesis with original data, bootstrap, high iterations and a low cp

syn_nb_hcp = synthesize(imputed_data, nsim = 1000) #synthesis with imputed data with no bootstrap and a high cp
syn_nb_lcp = synthesize(imputed_data, nsim = 1000, cp = 1e-32) #synthesis with imputed data with no bootstrap and a low cp
hi_b_hcp = synthesize(imputed_data, nsim = 1000, bootstrap = TRUE, maxit = 5)  #synthesis with imputed data with a high amount of iterations, bootstrap and a high cp
hi_b_lcp = synthesize(imputed_data, nsim = 1000, bootstrap = TRUE, cp = 1e-32, maxit = 5) #synthesis with imputed data with a high amount of iterations, bootstrap and a low cp
li_b_hcp = synthesize(imputed_data, nsim = 1000, bootstrap = TRUE)  #synthesis with imputed data with a low amount of iterations, bootstrap and a high cp
li_b_lcp = synthesize(imputed_data, nsim = 1000, bootstrap = TRUE, cp = 1e-32) #synthesis with imputed data with a low amount of iterations, bootstrap and a low cp
```

## 5b. Calculating the statistical properties of the synthetic sets

```{r, warning=FALSE}
all_synthetic_statistics = data.frame()
for (i in seq(1, length(synthesized), 1)) { #looping through each simulation
  imputation = synthesized[[i]]
  for (n in seq(1, imputation$m, 1)) { #looping through each imputation of the simulation
    syndata = complete(imputation, action = n)
    synthetic_statistics = statistical_properties(syndata)
    
    all_synthetic_statistics = rbind(all_synthetic_statistics, synthetic_statistics)
  }
}
```

## 5c. Running the analysis model on each of the synthetic sets

```{r}
all_synthetic_parameters = data.frame()
for (i in seq(1, length(synthesized), 1)) { #looping through each simulation
  imputation = synthesized[[i]]
  for (n in seq(1, imputation$m, 1)) { #looping through each imputation of the simulation
    syndata = complete(imputation, action = n)
    synthetic_parameters = glm(Outcome~BMI+Glucose+Pregnancies, data = syndata, family = "binomial") %>% tidy()
    all_synthetic_parameters = rbind(all_synthetic_parameters, synthetic_parameters)
  }
}
```

## 5d. Combining the statistics and parameters into single inferences

```{r, warning=FALSE}
single_synthetic_statistics = data.frame()
for (i in seq(1, length(unique(all_synthetic_statistics$term)))) {
  term_data = all_synthetic_statistics[all_synthetic_statistics$term == unique(all_synthetic_statistics$term)[i],] %>%
    sapply(mean)
  term_data$term = unique(all_synthetic_statistics$term)[i]
  term_data = data.frame(term_data)
  single_synthetic_statistics = rbind(single_synthetic_statistics, term_data)
}

single_synthetic_parameters = data.frame()
for (i in seq(1, length(unique(all_synthetic_parameters$term)))) {
  term_data = all_synthetic_parameters[all_synthetic_parameters$term == unique(all_synthetic_parameters$term)[i],] %>%
    sapply(mean)
  term_data$term = unique(all_synthetic_parameters$term)[i]
  term_data = data.frame(term_data)
  single_synthetic_parameters = rbind(single_synthetic_parameters, term_data)
}
```

## 5e & 5f. Combining half of the imputed set randomly with half of the true set for each of the m imputed sets & running a prediction model that predicts syn or true

```{r}
print(predict_synth_or_true(syn_nb_hcp, imputed_data))
print(predict_synth_or_true(synthesized, data))
```

## 5g. Pooling the model parameters

```{r}
pooled = map_dfr(synthesized, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled2 = map_dfr(ori_nb_lcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled3 = map_dfr(orl_b_hcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled3norm = map_dfr(orl_b_hcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool.normal()
})

pooled4 = map_dfr(orl_b_lcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled4norm = map_dfr(orl_b_lcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool.normal()
})

pooled5 = map_dfr(orh_b_hcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled5norm = map_dfr(orh_b_hcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool.normal()
})

pooled6 = map_dfr(orh_b_lcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled6norm = map_dfr(orh_b_lcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool.normal()
})

pooled7 = map_dfr(syn_nb_hcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled8 = map_dfr(syn_nb_lcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled9 = map_dfr(hi_b_hcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled9norm = map_dfr(hi_b_hcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool.normal()
})

pooled10 = map_dfr(hi_b_lcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled10norm = map_dfr(hi_b_lcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool.normal()
})

pooled11 = map_dfr(li_b_hcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled11norm = map_dfr(li_b_hcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool.normal()
})

pooled12 = map_dfr(li_b_lcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool3.syn()
})

pooled12norm = map_dfr(li_b_lcp, function(x) {
  x %$%
   glm(Outcome ~ BMI+Glucose+Pregnancies, family = 'binomial') %>%
  pool.normal()
})
```

## 6a. Comparing the synthetic statistics and parameters with the true statistics and parameters

```{r}
true_parameters = data.frame(tidy(analysis_model))
statistics_difference = single_synthetic_statistics[,2:ncol(single_synthetic_statistics)] - true_statistics[,2:ncol(true_statistics)]
parameters_difference = single_synthetic_parameters[,2:ncol(single_synthetic_parameters)] - true_parameters[,2:ncol(true_parameters)]
statistics_difference$term = single_synthetic_statistics$term
parameters_difference$term = single_synthetic_parameters$term
statistics_difference = statistics_difference[,c(ncol(statistics_difference),1:(ncol(statistics_difference)-1))]
parameters_difference = parameters_difference[,c(ncol(parameters_difference),1:(ncol(parameters_difference)-1))]

print(statistics_difference)
print(parameters_difference)
```

## 6b. Plotting the distributions of the statistics and parameters

```{r, warning=FALSE}
repeat_distribution = function(dataframe) {
  for (i in seq(1, length(unique(dataframe$term)), 1)) {
    print(unique(dataframe$term)[i])
    print(dataset_distribution(dataframe[dataframe$term == unique(dataframe$term)[i],]))
  }
}

print("Statistics")
repeat_distribution(all_synthetic_statistics)
print("Parameters")
repeat_distribution(all_synthetic_parameters)
print("Pooling parameters")
repeat_distribution(pooled)
```

## 6c. Calculating the confidence interval coverages

```{r}
true_data = glm(Outcome~BMI+Glucose+Pregnancies, data = data, family = "binomial")
true_data_imputed = glm(Outcome~BMI+Glucose+Pregnancies, data = imputed_data, family = "binomial")


confidence_interval = ci_cov(pooled, true_fit = true_data)
confidence_interval2 = ci_cov(pooled2, true_fit = true_data)
confidence_interval3 = ci_cov(pooled3, true_fit = true_data)
confidence_interval4 = ci_cov(pooled4, true_fit = true_data)
confidence_interval5 = ci_cov(pooled5, true_fit = true_data)
confidence_interval6 = ci_cov(pooled6, true_fit = true_data)
confidence_interval3norm = ci_cov(pooled3norm, true_fit = true_data)
confidence_interval4norm = ci_cov(pooled4norm, true_fit = true_data)
confidence_interval5norm = ci_cov(pooled5norm, true_fit = true_data)
confidence_interval6norm = ci_cov(pooled6norm, true_fit = true_data)
confidence_interval7 = ci_cov(pooled7, true_fit = true_data_imputed)
confidence_interval8 = ci_cov(pooled8, true_fit = true_data_imputed)
confidence_interval9 = ci_cov(pooled9, true_fit = true_data_imputed)
confidence_interval10 = ci_cov(pooled10, true_fit = true_data_imputed)
confidence_interval11 = ci_cov(pooled11, true_fit = true_data_imputed)
confidence_interval12 = ci_cov(pooled12, true_fit = true_data_imputed)
confidence_interval9norm = ci_cov(pooled9norm, true_fit = true_data_imputed)
confidence_interval10norm = ci_cov(pooled10norm, true_fit = true_data_imputed)
confidence_interval11norm = ci_cov(pooled11norm, true_fit = true_data_imputed)
confidence_interval12norm = ci_cov(pooled12norm, true_fit = true_data_imputed)

# print(xtable(confidence_interval[,c(1, 4, 8:11)]))
# print(xtable(confidence_interval2[,c(1, 4, 8:11)]))
# print(xtable(confidence_interval3[,c(1, 4, 8:11)]))
# print(xtable(confidence_interval4[,c(1, 4, 8:11)]))
# print(xtable(confidence_interval5[,c(1, 4, 8:11)]))
# print(xtable(confidence_interval6[,c(1, 4, 8:11)]))
# print(xtable(confidence_interval7[,c(1, 4, 8:11)]))
# print(xtable(confidence_interval8[,c(1, 4, 8:11)]))
# print(xtable(confidence_interval9[,c(1, 4, 8:11)]))
# print(xtable(confidence_interval10[,c(1, 4, 8:11)]))
# print(xtable(confidence_interval11[,c(1, 4, 8:11)]))
# print(xtable(confidence_interval12[,c(1, 4, 8:11)]))
```

```{r}
print("The standard synthesis with original data, no bootstrap and a 1e-4 cp")
print(confidence_interval[,c(1, 4, 8:11)])
print("Synthesis with original data, no bootstrap and a 1e-32 cp")
print(confidence_interval2[,c(1, 4, 8:11)])
print("Synthesis with original data, bootstrap, 1 iteration and a 1e-4 cp")
print(confidence_interval3[,c(1, 4, 8:11)])
print("Synthesis with original data, bootstrap, 1 iteration and a 1e-32 cp")
print(confidence_interval4[,c(1, 4, 8:11)])
print("synthesis with original data, bootstrap, 5 iterations and a 1e-4 cp")
print(confidence_interval5[,c(1, 4, 8:11)])
print("synthesis with original data, bootstrap, 5 iterations and a 1e-32 cp")
print(confidence_interval6[,c(1, 4, 8:11)])

print("Synthesis with original data, bootstrap, 1 iteration and a 1e-4 cp with pool.normal")
print(confidence_interval3norm[,c(1, 4, 8:11)])
print("Synthesis with original data, bootstrap, 1 iteration and a 1e-32 cp with pool.normal")
print(confidence_interval4norm[,c(1, 4, 8:11)])
print("synthesis with original data, bootstrap, 5 iterations and a 1e-4 cp with pool.normal")
print(confidence_interval5norm[,c(1, 4, 8:11)])
print("synthesis with original data, bootstrap, 5 iterations and a 1e-32 cp with pool.normal")
print(confidence_interval6norm[,c(1, 4, 8:11)])

print("synthesis with imputed data with no bootstrap and a 1e-4 cp")
print(confidence_interval7[,c(1, 4, 8:11)])
print("synthesis with imputed data with no bootstrap and a 1e-32 cp")
print(confidence_interval8[,c(1, 4, 8:11)])
print("synthesis with imputed data with 5 iterations, bootstrap and a 1e-4 cp")
print(confidence_interval9[,c(1, 4, 8:11)])
print("synthesis with imputed data with 5 iterations, bootstrap and a 1e-32 cp")
print(confidence_interval10[,c(1, 4, 8:11)])
print("synthesis with imputed data with 1 iteration, bootstrap and a 1e-4 cp")
print(confidence_interval11[,c(1, 4, 8:11)])
print("synthesis with imputed data with 1 iteration, bootstrap and a 1e-32 cp")
print(confidence_interval12[,c(1, 4, 8:11)])

print("synthesis with imputed data with 5 iterations, bootstrap and a 1e-4 cp with pool.normal")
print(confidence_interval9norm[,c(1, 4, 8:11)])
print("synthesis with imputed data with 5 iterations, bootstrap and a 1e-32 cp with pool.normal")
print(confidence_interval10norm[,c(1, 4, 8:11)])
print("synthesis with imputed data with 1 iteration, bootstrap and a 1e-4 cp with pool.normal")
print(confidence_interval11norm[,c(1, 4, 8:11)])
print("synthesis with imputed data with 1 iteration, bootstrap and a 1e-32 cp with pool.normal")
print(confidence_interval12norm[,c(1, 4, 8:11)])
```

```{r}
sessionInfo()
```

